---
title: "MOON_analysis"
author: "Anne-Claire Kroger"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(decoupleR)
library(OmnipathR)
library(readr)
library(cosmosR)
library(MOFA2)
library(stringr)
library(data.table)
library(pheatmap)
library(ggplot2)
```


#Import data
```{r}

#MOFA results 
gbm_MOFAobject <- readRDS("mofa_results/mofa_gbm_prot_phospho_f15.RDS")
gbm_MOFAobject_REG <- readRDS("mofa_results/mofa_gbm_NOprot_REGphospho_f15.RDS")

brca_MOFAobject <- readRDS("mofa_results/mofa_brca_prot_phospho_f15.RDS")
brca_MOFAobject_REG <- readRDS("mofa_results/mofa_brca_NOprot_REGphospho_f15.RDS")

```
Everything with "REG" in the name means the proteomics abundance was regressed out of the phosphoproteomics data before running MOFA.

```{R}
#Investigate MOFA factors 
MOFAobject <- brca_MOFAobject_REG

pheatmap(MOFAobject@cache$variance_explained$r2_per_factor[[1]], display_numbers = T, angle_col = "0", legend_labels = c("0","10", "20", "30", "40", "Variance\n\n"), legend = T, main = "", legend_breaks = c(0,10, 20, 30, 40, max(MOFAobject@cache$variance_explained$r2_per_factor[[1]])), cluster_rows = F, cluster_cols = F, color = colorRampPalette(c("white","red"))(100), fontsize_number = 10)

correlate_factors_with_covariates(MOFAobject, 
                                    covariates = c("PAM50", "NMF.Cluster"))
  
plot_factor(MOFAobject, 
            factors = c(1,3,4),        #change factor you want to display
            #color_by = "Proteome_2023",
            #color_by = "PAM50",
            color_by = "NMF.Cluster",
            #color_by = "multiomic",
            #color_by = "rna_wang_cancer_cell_2017",
            dot_size = 3,        
            dodge = T,           
            legend = T,          
            add_boxplot = T  
)

```
Factors of interest

GBM: Factor 1,2,5(IDH),(3,7)
GBM_REG: Factor 1,2,6(IDH),(5)

BRCA: Factor 1,2,4 
BRCA_REG: 1,4

```{r}
#brca
brca_weights_rna <- get_weights(brca_MOFAobject, factors = "all")$Transcriptome
brca_weights_phospho <- get_weights(brca_MOFAobject, factors = "all")$Phosphoproteome

rownames(brca_weights_rna) <- str_replace(rownames(brca_weights_rna), "_.*", "")

brca_weightsREG_rna <- get_weights(brca_MOFAobject_REG, factors = "all")$Transcriptome
brca_weightsREG_phospho <- get_weights(brca_MOFAobject_REG, factors = "all")$REG_Phosphoproteome

rownames(brca_weightsREG_rna) <- str_replace(rownames(brca_weightsREG_rna), "_.*", "")

#gbm
gbm_weights_rna <- get_weights(gbm_MOFAobject, factors = "all")$Transcriptome
gbm_weights_phospho <- get_weights(gbm_MOFAobject, factors = "all")$Phosphoproteome

rownames(gbm_weights_rna) <- str_replace(rownames(gbm_weights_rna), "_.*", "")

gbm_weightsREG_rna <- get_weights(gbm_MOFAobject_REG, factors = "all")$Transcriptome
gbm_weightsREG_phospho <- get_weights(gbm_MOFAobject_REG, factors = "all")$REG_Phosphoproteome

rownames(gbm_weightsREG_rna) <- str_replace(rownames(gbm_weightsREG_rna), "_.*", "")

#scaling not necessary because we are only using one data type for upstream nodes
```

#Do basic TF activity and Kinase activity prediction on all MOFA factors

##Get target networks: TF targets, Kinase targets
```{r}
#Kinase targets (code from aurelien old script)
KSN <- OmnipathR::import_omnipath_enzsub()
KSN <- KSN[!KSN$sources == "KEA",]
KSN$target <- paste(KSN$substrate_genesymbol, paste(KSN$residue_type, KSN$residue_offset, sep = ""), sep = "_")
KSN$mor <- ifelse(KSN$modification == "phosphorylation", 1, ifelse(KSN$modification == "dephosphorylation", -1, NA))
KSN <- KSN[,c(3,13,14)]
names(KSN) <- c("source","target","mor")
KSN$enzsub <- paste(KSN$source, KSN$target, sep = "_")
dubs <- KSN[duplicated(KSN$enzsub), "enzsub", drop = T]
KSN <- KSN[-which(KSN$enzsub %in% dubs),]
KSN <- KSN[,-4]

dim(KSN) #ressource has 36495 commections with 670 Kinase (sources) with at least 5 connections

saveRDS(KSN, "network_ressources/KSN_omnipath_noKEA_network.RDS")
```

```{r}
#Load Kinase targets
KSN <- readRDS("network_ressources/KSN_omnipath_noKEA_network.RDS")

#Load TF targets (from Aurelien MOON script)
TFN <- decoupleR::get_collectri()

```


##TF activity prediction on Transcriptomics layer of MOFA results with ulm
```{r}
#brca
brca_tf <- decoupleR::run_ulm(brca_weights_rna, 
                              TFN, 
                              .source='source', 
                              .target='target', 
                              minsize = 5) 

brca_tf_df <- as.data.frame(dcast(as.data.table(brca_tf[,2:4]), source ~ condition, value.var = "score"))
rownames(brca_tf_df) <- brca_tf_df$source
brca_tf_df <- brca_tf_df[,-1]

saveRDS(brca_tf, "kin_tf_prediction_results/brca_mofa_tf_ulm.RDS")

brcaREG_tf <- decoupleR::run_ulm(brca_weightsREG_rna, 
                              TFN, 
                              .source='source', 
                              .target='target', 
                              minsize = 5) 

brcaREG_tf_df <- as.data.frame(dcast(as.data.table(brcaREG_tf[,2:4]), source ~ condition, value.var = "score"))
rownames(brcaREG_tf_df) <- brcaREG_tf_df$source
brcaREG_tf_df <- brcaREG_tf_df[,-1]

saveRDS(brcaREG_tf, "kin_tf_prediction_results/brca_mofaREG_tf_ulm.RDS")

#gbm
gbm_tf <- decoupleR::run_ulm(gbm_weights_rna, 
                              TFN, 
                              .source='source', 
                              .target='target', 
                              minsize = 5) 

gbm_tf_df <- as.data.frame(dcast(as.data.table(gbm_tf[,2:4]), source ~ condition, value.var = "score"))
rownames(gbm_tf_df) <- gbm_tf_df$source
gbm_tf_df <- gbm_tf_df[,-1]

saveRDS(gbm_tf, "kin_tf_prediction_results/gbm_mofa_tf_ulm.RDS")

gbmREG_tf <- decoupleR::run_ulm(gbm_weightsREG_rna, 
                              TFN, 
                              .source='source', 
                              .target='target', 
                              minsize = 5) 

gbmREG_tf_df <- as.data.frame(dcast(as.data.table(gbmREG_tf[,2:4]), source ~ condition, value.var = "score"))
rownames(gbmREG_tf_df) <- gbmREG_tf_df$source
gbmREG_tf_df <- gbmREG_tf_df[,-1]

saveRDS(gbmREG_tf, "kin_tf_prediction_results/gbm_mofaREG_tf_ulm.RDS")

```


##Kinase activity prediction on Phosphoproteomics layer of MOFA results with ulm
```{r}
#brca
brca_kin <- decoupleR::run_ulm(brca_weights_phospho, 
                              KSN, 
                              .source='source', 
                              .target='target', 
                              minsize = 3) 

brca_kin_df <- as.data.frame(dcast(as.data.table(brca_kin[,2:4]), source ~ condition, value.var = "score"))
rownames(brca_kin_df) <- brca_kin_df$source
brca_kin_df <- brca_kin_df[,-1]

saveRDS(brca_kin, "kin_tf_prediction_results/brca_mofa_kin_ulm.RDS")

brcaREG_kin <- decoupleR::run_ulm(brca_weightsREG_phospho, 
                              KSN, 
                              .source='source', 
                              .target='target', 
                              minsize = 3) 

brcaREG_kin_df <- as.data.frame(dcast(as.data.table(brcaREG_kin[,2:4]), source ~ condition, value.var = "score"))
rownames(brcaREG_kin_df) <- brcaREG_kin_df$source
brcaREG_kin_df <- brcaREG_kin_df[,-1]

saveRDS(brcaREG_kin, "kin_tf_prediction_results/brca_mofaREG_kin_ulm.RDS")

#gbm
gbm_kin <- decoupleR::run_ulm(gbm_weights_phospho, 
                              KSN, 
                              .source='source', 
                              .target='target', 
                              minsize = 3) 

gbm_kin_df <- as.data.frame(dcast(as.data.table(gbm_kin[,2:4]), source ~ condition, value.var = "score"))
rownames(gbm_kin_df) <- gbm_kin_df$source
gbm_kin_df <- gbm_kin_df[,-1]

saveRDS(gbm_kin, "kin_tf_prediction_results/gbm_mofa_kin_ulm.RDS")

gbmREG_kin <- decoupleR::run_ulm(gbm_weightsREG_phospho, 
                              KSN, 
                              .source='source', 
                              .target='target', 
                              minsize = 3) 

gbmREG_kin_df <- as.data.frame(dcast(as.data.table(gbmREG_kin[,2:4]), source ~ condition, value.var = "score"))
rownames(gbmREG_kin_df) <- gbmREG_kin_df$source
gbmREG_kin_df <- gbmREG_kin_df[,-1]

saveRDS(gbmREG_kin, "kin_tf_prediction_results/gbm_mofaREG_kin_ulm.RDS")
```


##Load all TF and Kinase activity prediction results
```{r}
brca_tf <- readRDS("kin_tf_prediction_results/brca_mofa_tf_ulm.RDS")
brcaREG_tf <- readRDS("kin_tf_prediction_results/brca_mofaREG_tf_ulm.RDS")

gbm_tf <- readRDS("kin_tf_prediction_results/gbm_mofa_tf_ulm.RDS")
gbmREG_tf <- readRDS("kin_tf_prediction_results/gbm_mofaREG_tf_ulm.RDS")

brca_kin <- readRDS("kin_tf_prediction_results/brca_mofa_kin_ulm.RDS")
brcaREG_kin <- readRDS("kin_tf_prediction_results/brca_mofaREG_kin_ulm.RDS")

gbm_kin <- readRDS("kin_tf_prediction_results/gbm_mofa_kin_ulm.RDS")
gbmREG_kin <- readRDS("kin_tf_prediction_results/gbm_mofaREG_kin_ulm.RDS")
```


#Run MOON Kinase activty prediction based on TF activity prediction

##Prep the Kinase-TF interactions network from omnipath
```{r}
#Pull base network from Omnipath
full_pkn <- as.data.frame(import_omnipath_interactions())

#Clean connections 
full_pkn <- full_pkn[!is.na(full_pkn$references),]
clean_PKN <- full_pkn[full_pkn$consensus_stimulation == 1 | full_pkn$consensus_inhibition == 1,]

#Define sign 
clean_PKN$sign <- clean_PKN$consensus_stimulation - clean_PKN$consensus_inhibition
clean_PKN <- clean_PKN[,c(3,4,16)]

#Repeat interactions with sign 0 to be in the list with both -1 and 1 sign
clean_PKN_supp <- clean_PKN[clean_PKN$sign == 0,]
clean_PKN_supp$sign <- -1
clean_PKN[clean_PKN$sign == 0,"sign"] <- 1
clean_PKN <- as.data.frame(rbind(clean_PKN, clean_PKN_supp))

#Final format
names(clean_PKN) <- c("source","target","interaction")
clean_PKN <- clean_PKN[,c(1,3,2)]

#Final clean
meta_network_filtered <- meta_network_cleanup(clean_PKN)

write_csv(x = clean_PKN, file = "network_ressources/clean_omnipath_PKN.csv")
```

##Filter & compress the network according to our input TFs
```{r}
#Choose Factor and data set and run the whole chunk
#TF_data <- gbmREG_tf
#RNA_data <- gbm_weightsREG_rna
#kin_data <- gbmREG_kin
#factor_n <- 6

#TF_data <- brcaREG_tf
#RNA_data <- brca_weightsREG_rna
#kin_data <- brcaREG_kin
#factor_n <- 4

TF_data <- gbm_tf
RNA_data <- gbm_weights_rna
kin_data <- gbm_kin
factor_n <- 5

n_steps <- 6


clean_PKN <- read_csv("network_ressources/clean_omnipath_PKN.csv")

#RNA input for later filtering
RNA_input <- RNA_data[,factor_n]
names(RNA_input) <- rownames(RNA_data)

#TF input from MOFA as downstream input
TF_inputs_f <- TF_data[TF_data$condition == paste0("Factor", factor_n),]

#TF results for one factor as named vector
TF_inputs <- TF_inputs_f$score
names(TF_inputs) <- TF_inputs_f$source
downstream_inputs <- TF_inputs

#filter TFs (downstream input) for TFs actually in the PKN
downstream_inputs_filtered <- cosmosR:::filter_input_nodes_not_in_pkn(downstream_inputs, 
                                                                      meta_network_filtered)

#filter PKN for TFs that are downstream of our Kinases we have in the network 
meta_network_filtered <- cosmosR:::keep_observable_neighbours(meta_network_filtered, 
                                                              n_steps, 
                                                              names(downstream_inputs_filtered))

#compress PKN in a way that nodes with the same children are united
meta_network_compressed_list <- compress_same_children(meta_network_filtered, 
                                                       sig_input = c("wig"=""), #give it an empty name vector
                                                       metab_input = downstream_inputs_filtered)
meta_network_compressed <- meta_network_compressed_list$compressed_network

#cleanup again to remove empty/ redunant connections
meta_network_compressed <- meta_network_cleanup(meta_network_compressed)

```

##Run MOON in interation
```{r}
meta_network_rec_to_TFmetab <- meta_network_compressed

before <- 1
after <- 0
i <- 1
while (before != after & i < 10) {
  before <- length(meta_network_rec_to_TFmetab[,1])
  recursive_decoupleRnival_res <- cosmosR::moon( downstream_input = downstream_inputs_filtered, 
                                                 meta_network = meta_network_rec_to_TFmetab, 
                                                 n_layers = n_steps, 
                                                 statistic = "ulm") 
  
  meta_network_rec_to_TFmetab <- filter_incohrent_TF_target(recursive_decoupleRnival_res, TFN, meta_network_rec_to_TFmetab, RNA_input)
  after <- length(meta_network_rec_to_TFmetab[,1])
  i <- i + 1
}

```

#Compare and save results

##Build comparason table
```{r}
kin_results <- kin_data 

results_moon <- recursive_decoupleRnival_res
all_results <- kin_results[kin_results$condition == paste0("Factor", factor_n), c(2,4)]

all_results$moon_score <- results_moon$score[match(all_results$source, results_moon$source)]
all_results$moon_level <- as.factor(results_moon$level[match(all_results$source, results_moon$source)])
```

##Visualization
```{r}
ggplot(all_results, aes(x=moon_score, y=score, color=moon_level)) +
  geom_point()

reg <- lm(score~moon_score, data = all_results[all_results$moon_level==1,])
summary(reg)

temp <- all_results[all_results$moon_level == 1,]

ggplot(temp, aes(x=moon_score, y=score, color=moon_level)) +
  geom_point()

cor(temp$moon_score, temp$score, use = "pairwise.complete.obs", method = "spearman")

```
## Save all results
```{r}

#gbmf1_moon <- results_moon 
#gbmf1_comparison <- all_results

#gbmf2_moon <- results_moon 
#gbmf2_comparison <- all_results


#gbmf5_moon <- results_moon 
#gbmf5_comparison <- all_results

#gbmf1_REG_moon <- results_moon 
#gbmf1_REG_comparison <- all_results

#gbmf2_REG_moon <- results_moon 
#gbmf2_REG_comparison <- all_results

#This one has a very strong biological factor but unfortunately no good moon predictions 
#gbmf6_REG_moon <- results_moon 
#gbmf6_REG_comparison <- all_results
```
Best results are wth GBM_REG & GBM factor 1.


```{r}

#brcaf1_moon <- results_moon 
#brcaf1_comparison <- all_results

#brcaf1_REG_moon <- results_moon 
#brcaf1_REG_comparison <- all_results

#brcaf4_REG_moon <- results_moon 
#brcaf4_REG_comparison <- all_results

```
BRCA overall looks worse then GBM...

```{r}
save(gbmf1_moon, gbmf1_comparison, 
     gbmf2_moon, gbmf2_comparison, 
     gbmf5_moon, gbmf5_comparison,
     gbmf1_REG_moon, gbmf1_REG_comparison,
     gbmf2_REG_moon, gbmf2_REG_comparison,
     gbmf6_REG_moon, gbmf6_REG_comparison,
     brcaf1_moon, brcaf1_comparison,
     brcaf1_REG_moon, brcaf1_REG_comparison,
     brcaf4_REG_moon, brcaf4_REG_comparison,
     file = "cptac_moon_results.RData")
```

##Load all results
```{r}
load("cptac_moon_results.RData")

results_moon <- gbmf1_REG_moon 
all_results <- gbmf1_REG_comparison
```

